---
title: "Biostat M280 Homework 1"
author: "Xinrui Zhang UID:605231530"
subtitle: Due Jan 25 @ 11:59PM
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1. Git/GitHub

**No handwritten homework reports are accepted for this course.** We work with Git and GitHub. Efficient and abundant use of Git, e.g., frequent and well-documented commits, is an important criterion for grading your homework.

1. Apply for the [Student Developer Pack](https://education.github.com/pack) at GitHub using your UCLA email.

2. Create a **private** repository `biostat-m280-2019-winter` and add `Hua-Zhou` and `juhkim111` as your collaborators with write permission.

3. Top directories of the repository should be `hw1`, `hw2`, ... Maintain two branches `master` and `develop`. The `develop` branch will be your main playground, the place where you develop solution (code) to homework problems and write up report. The `master` branch will be your presentation area. Submit your homework files (R markdown file `Rmd`, `html` file converted from R markdown, all code and data sets to reproduce results) in `master` branch.

4. After each homework due date, teaching assistant and instructor will check out your master branch for grading. Tag each of your homework submissions with tag names `hw1`, `hw2`, ... Tagging time will be used as your submission time. That means if you tag your `hw1` submission after deadline, penalty points will be deducted for late submission.

## Q2. Linux Shell Commands

The `/home/m280data/NYCHVS` folder on teaching server contains a data set from the New York City Housing and Vacancy Survey. See [2019 ASA Data Challenge Expo](https://www1.nyc.gov/site/hpd/about/nychvs-asa-data-challenge-expo.page) for further details.  
    ```{bash}
    ls -l /home/m280data/NYCHVS
    ```

1. Display the first few lines of `NYCHVS_1991.csv`.

    **SOLUTION**
    ```{bash}
    head -4 /home/m280data/NYCHVS/NYCHVS_1991.csv
    ```


2. Display number of lines in each `csv` file.

    **SOLUTION**
    ```{bash}
    wc -l /home/m280data/NYCHVS/*.csv
    ```


3. Display the 3 files that have the least number of lines

    **SOLUTION**
    ```{bash}
    wc -l /home/m280data/NYCHVS/*.csv | sort | head -3
    ```


4. What's the output of following bash script?
    ```{bash, eval=FALSE}
    for datafile in /home/m280data/NYCHVS/*.csv
      do
        ls $datafile
      done
    ```

    **SOLUTION**

    It displays all csv files in the directory `/home/m280data/NYCHVS` without changing your current directory

5. What unique values does the second variable `borough` take in `NYCHVS_1991.csv`? Tabulate how many times each value appears.

    **SOLUTION**
    
    ```{bash}
    # to see which coloumn variable 'borough' situates in
    grep borough /home/m280data/NYCHVS/NYCHVS_1991.csv

    # 'uniq -c' shows how many times one element occurs in the file
    awk -F, '{print$2}' /home/m280data/NYCHVS/NYCHVS_1991.csv | sort | uniq -c | head -5
    ```
    Thus `borough` takes 1,2,3,4,5 as its unique values 

## Q3. More fun with shell

1. You and your friend just have finished reading *Pride and Prejudice* by Jane Austen. Among the four main characters in the book, Elizabeth, Jane, Lydia, and Darcy, your friend thinks that Darcy was the most mentioned. You, however, are certain it was Elizabeth. Obtain the full text of the novel from <https://www.gutenberg.org/files/1342/1342.txt> and save to your local folder. 
    ```{bash, eval=FALSE}
    curl https://www.gutenberg.org/files/1342/1342.txt > pride_and_prejudice.txt
    ```

    **SOLUTION**
    ```{bash}
    for namei in Elizabeth Jane Lydia Darcy
     do
      echo $namei
      grep -o $namei /home/suedez/pride_and_prejudice.txt | wc -l
      echo
     done
    ```


2. What's the difference between the following two commands?
    ```{bash eval=FALSE}
    echo 'hello, world' > test1.txt
    ```
    and
    ```{bash eval=FALSE}
    echo 'hello, world' >> test2.txt
    ```

    **SOLUTION**
  
    The first one write `hello, world` into a text file by covering the old contents of that text file, while the second one just append `hello, world` to the text file without changing old contents.

3. Using your favorite text editor (e.g., `vi`), type the following and save the file as `middle.sh`:
    ```{bash eval=FALSE}
    #!/bin/sh
    # Select lines from the middle of a file.
    # Usage: bash middle.sh filename end_line num_lines
    head -n "$2" "$1" | tail -n "$3"
    ```

    **SOLUTION**
  
    1) After edit, let't look at the file `middle.sh`
    ```{bash}
    cat middle.sh
    ```
   
    2) Using `chmod` make the file executable by the owner, and run 
    ```{bash eval=TRUE}
    chmod u+x middle.sh
    ./middle.sh pride_and_prejudice.txt 20 5
    ```
    
    3) Explain the output. Explain the meaning of `"$1"`, `"$2"`, and `"$3"` in this shell script. Why do we need the first line of the shell script?

    `"$2"`means take the second input as the input for `n` of `head -n`; `"$1"` means take the first input as the   input for `filename`; `"$3"` means take the third input as input for `n` of `tail -n`.
    
    The first line tells the system to use `/bin/sh` as the interprater to run the following script.



## Q4. R Batch Run

In class we discussed using R to organize simulation studies. 

1. Expand the [`runSim.R`](http://hua-zhou.github.io/teaching/biostatm280-2019winter/slides/02-linux/runSim.R) script to include arguments `seed` (random seed), `n` (sample size), `dist` (distribution) and `rep` (number of simulation replicates). When `dist="gaussian"`, generate data from standard normal; when `dist="t1"`, generate data from t-distribution with degree of freedom 1 (same as Cauchy distribution); when `dist="t5"`, generate data from t-distribution with degree of freedom 5. Calling `runSim.R` will (1) set random seed according to argument `seed`, (2) generate data according to argument `dist`, (3) compute the primed-indexed average estimator and the classical sample average estimator for each simulation replicate, (4) report the average mean squared error (MSE)
$$
  \frac{\sum_{r=1}^{\text{rep}} (\widehat \mu_r - \mu_{\text{true}})^2}{\text{rep}}
$$
for both methods.

    **SOLUTION**
    
    1) Expand the `runSim.R` as it required, and let's have a look at the expanded `runSim.R`
    ```{bash}
    cat runSim.R
    ```
    
     2) Run the R code with certain arguments passed from linux
    ```{bash}
    Rscript runSim.R  n=100 dist=\"gaussian\" rep=50 seed=280 # pass string to R
    ```
  

2. Modify the [`autoSim.R`](http://hua-zhou.github.io/teaching/biostatm280-2019winter/slides/02-linux/autoSim.R) script to run simulations with combinations of sample sizes `nVals = seq(100, 500, by=100)` and distributions `distTypes = c("gaussian", "t1", "t5")` and write output to appropriately named files. Use `rep = 50`, and `seed = 280`. 

    **SOLUTION**

    1. Modify the `autoSim.R` to meet the requirements and have a look on that R scripts
    ```{bash}
    cat autoSim.R
    ```
    
    2. Run the R code 
    ```{bash}
    Rscript autoSim.R seed=280 rep=50
    ```
  
3. Write an R script to collect simulation results from output files and print average MSEs in a table of format

| $n$ | Method   | Gaussian | $t_5$ | $t_1$ |
|-----|----------|-------|-------|----------|
| 100 | PrimeAvg |       |       |          |
| 100 | SampAvg  |       |       |          |
| 200 | PrimeAvg |       |       |          |
| 200 | SampAvg  |       |       |          |
| 300 | PrimeAvg |       |       |          |
| 300 | SampAvg  |       |       |          |
| 400 | PrimeAvg |       |       |          |
| 400 | SampAvg  |       |       |          |
| 500 | PrimeAvg |       |       |          |
| 500 | SampAvg  |       |       |          |

    
    **SOLUTION**
    
```{r}
    ## read MSEs from txt files and built a vector to save them
    nVals <- seq(100, 500, by=100)
    distTypes = c("gaussian", "t1", "t5")
    results <- c() 
    N <- c() # built for the first coloumn of the table
    for (dist in distTypes) {
      for (n in nVals){
        oFile <- paste("/home/suedez/biostat-m280-2019-winter/hw1/n",n,
                       "dist",dist, ".txt", sep="")
        Tresults <- read.table(oFile)
        Tresults <- Tresults[,2]
        results <- c(results,Tresults)
       N <- c(N,rep(n,2))
      }
    }

    ## built the 2nd coloumn for the table
    Method <- rep(c('PrimeAvg','SampAvg'),5)

    ## built a table to display results
    results_table <- data.frame(n=N,Method=Method,Gaussian=results[1:10],t_1=results[11:20],
                            t_5=results[21:30])
    results_table
```



