---
title: "Biostat M280 Homework 3"
subtitle: Due Mar 1 @ 11:59PM
author: Xinrui Zhang UID:605231530
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use tidyverse and bash to explore following two data sets.

## Q1 LA City Employee Payroll

The `/home/m280-data/la_payroll/City_Employee_Payroll.csv` file on teaching server contains payroll information of LA City employees in years 2013-2018. It was downloaded from [LA City Controller's Office](https://controllerdata.lacity.org/Payroll/City-Employee-Payroll/pazn-qyym). Make a Shiny app to facilitate visualization of this data. 

1. For efficiency of the Shiny app, you should first pre-process, pare down, tidy, and save the data, e.g., as a compressed RDS file, to be used in the app.

```{bash}
ln -sf /home/m280data/la_payroll/City_Employee_Payroll.csv LAEmployeePayroll.csv
ls -l NYParkingViolations.csv
```

####A.Load packages and read in cvs file
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
library(data.table)
library(dplyr)
library(rio)
library(magrittr)
city_ep <- read_csv('LAEmployeePayroll.csv') #city employee and payroll
colnames(city_ep)
```

####B.Process colnames and chose variables needed then export tibble in a rds file.

```{r message=FALSE, warning=FALSE}
names(city_ep) <- str_replace_all(names(city_ep), ".\\(.*\\)", "") 
#remove parenthesse
names(city_ep) <- str_replace_all(names(city_ep), " ", "_")
city_ep %<>%
  select(Department_Title, Job_Class_Title,
        Year, Total_Payments, Base_Pay,
        Overtime_Pay, Other_Pay,
        Average_Benefit_Cost)
##save as rds
library(rio)
export(city_ep, "LAep/LAep.rds")
```

0. Visualize any other information you are interested in.

I'd like to see the 6-year average payments composition of a chosen department.

0. Publish your Shiny app to <https://www.shinyapps.io> and share the link.

See my Shiny app on <https://suedez.shinyapps.io/laep/>.

## Q2 LA City Parking War

The SQLite database `/home/m280data/la_parking/LA_Parking_Citations.sqlite` on teaching server contains information about parking tickets in LA City. It was downloaded from [LA Open Data Portal](https://data.lacity.org/A-Well-Run-City/Parking-Citations/wjz9-h9np). Connect to the database and answer following questions using plots and summary statistics. In this exercise, you are **not** allowed to load whole data into memory. Use the _transform in database, plot in R_ strategy.

1. How many tickets are in this data set? Which time period do these tickets span? Which years have most data?

```{r message=FALSE, warning=FALSE}
library("DBI")
library("RSQLite")
```


```{bash}
ln -sf /home/m280data/la_parking/LA_Parking_Citations.sqlite LAparking.sqlite
ls -l LAparking.sqlite
```

```{r message=FALSE, warning=FALSE}
if (Sys.info()[["sysname"]] == "Linux") {
  db <- dbConnect(RSQLite::SQLite(), 
                  dbname = "LAparking.sqlite")
} else if (Sys.info()[["sysname"]] == "Darwin") {
  db <- dbConnect(RSQLite::SQLite(), 
                  dbname = "./LA_Parking_Citations.sqlite")
                  }
dbListTables(db)
LA_sql <- dplyr::tbl(db, "latix") #use tbl function in dplyr package
str(LA_sql)
```

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
LA_sql %>% summarise(n = n())
```
There are `7656418` tickets in total.


```{r message=FALSE, warning=FALSE}
##time range
options(scipen=100) # don't round
LA_sql %>% mutate(time = Issue_Year * 10e7 + 
                    Issue_Month * 10e5 + 
                    Issue_Day * 10e3 +
                    Issue_Hour * 10e1 + 
                    Issue_Minute) %>%
  summarise(Time_begin = min(time),
             Time_end = max(time))
##wich year most tickets
LA_sql %>% group_by(Issue_Year) %>% summarise(n = n()) %>% 
  ggplot() +
  geom_bar(stat = 'identity', aes(x = Issue_Year,
                                  y = n)) +
  labs(title = paste('Ammounts of tickets in year 2015-2019'), 
        x = 'Issue Year',
        y = 'Ammout of Tickets')
```

As it shown above, the time range is from `1:00,July 2nd, 2015` to `23:58, January 25th, 2019`. The year has most tickets is `2017`.

0. When (which hour, weekday, month day, and month) are you most likely to get a ticket and when are you least likely to get a ticket?

```{r message=FALSE, warning=FALSE}
LA_sql %>% group_by(Issue_Month) %>% summarise(n = n()) %>% 
  arrange(desc(n)) %>%
  select(Issue_Month, n) %>% 
  filter(is.na(Issue_Month) == F) %>%
  collect() -> MONTH
MONTH %>%
  filter(n == max(n) | n == min(n))
LA_sql %>% group_by(Issue_Day) %>% summarise(n = n()) %>% 
  arrange(desc(n)) %>%
  select(Issue_Day, n) %>% 
  filter(is.na(Issue_Day) == F) %>%
  collect() -> DAY
DAY %>%
  filter(n == max(n) | n == min(n)) 
LA_sql %>% group_by(Issue_Wday) %>% summarise(n = n()) %>% 
  arrange(desc(n)) %>%
  select(Issue_Wday, n) %>% 
  filter(is.na(Issue_Wday) == F) %>%
  collect() -> WDAY
WDAY %>%
  filter(n == max(n) | n == min(n))
LA_sql %>% group_by(Issue_Hour) %>% summarise(n = n()) %>% 
  arrange(desc(n)) %>%
  select(Issue_Hour, n) %>%
  filter(is.na(Issue_Hour) == F) %>%
  collect() -> HOUR
HOUR %>%
  filter(n == max(n) | n == min(n))


```

As most likely month, month day, week day, hour to get a ticket are, respectively, `August`, `13th`, `Wednesday`，`12:00`.
The least likely month, month day, week day, hour to get a ticket are, respectively, `February`, `31th`, `Monday`, `5:00`. 
    
0. Which car makes received most citations?

```{r message=FALSE, warning=FALSE}
LA_sql %>% group_by(Make) %>% summarise(n = n()) %>% 
  arrange(desc(n)) %>% filter(is.na(Make) == F) %>%
  select(Make, n)

```

`TOYT` received most citations.

0. How many different colors of cars were ticketed? Which color attracted most tickets?

```{r message=FALSE, warning=FALSE}
LA_sql %>% group_by(Color) %>% summarise(n = n()) %>% arrange(desc(n)) %>%
  filter(is.na(Color) == F) %>% collect()
```

There are 103 different colors.
The color `black` is the most attractive color, but maybe it is because black is the most common color for cars. 

0. What are the most common ticket types?

```{r message=FALSE, warning=FALSE}
LA_sql %>% group_by(Violation_Description) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n)) 
```

The most common ticket type is `No Park/Street Clean`.

0. How much money was collected on parking tickets in 2016, 2017 and 2018?

```{r message=FALSE, warning=FALSE}
LA_sql %>% filter(Issue_Year == 2016 | Issue_Year == 2017 |
                  Issue_Year == 2018) %>% select(Fine_amount, 
                                           Issue_Year) %>%
  group_by(Issue_Year) %>% 
  summarise(Money = sum(Fine_amount))  %>%
  ggplot(aes(x = Issue_Year, 
              y = Money, fill = Money)) + 
  geom_bar(stat = 'identity') +
  geom_text(aes(label = Money, 
                vjust = -0.1, 
                hjust = 0.5), 
            show_guide = FALSE,
            size = 3) +
  labs(title = 'Ammounts of fine collected in year 2016-2018', 
        x = 'Year',
        y = 'Ammout of Fine')
```

0. If you've been ticketed in LA County, did you find your ticket in this data set?

I don't even have a car hahaha.


0. Read the blog <http://www.brettrics.com/9-million-parking-tickets-la/> and try to reproduce plots using your data.

####a. Calendar Heatmap 
```{r message=FALSE, warning=FALSE}
brks <- seq(1, 12, 1)
lbs <- c('JAN', 'FEB', 'MAR', 'APR', 'MAY',
         'JUN', 'JUL', 'AUG', 'SEP',
         'OCT', 'NOV', 'DEC')
brksx <- seq(1, 31, 1)
lbsx <- 1:31
LA_sql %>% group_by(Issue_Year, Issue_Month, Issue_Day) %>%
  summarize(n = n()) %>% filter(is.na(Issue_Year) == F)  %>%
  mutate(Month = paste(Issue_Month)) %>%
  ggplot(aes(x = Issue_Day, y = Month,
         fill = n)) +
  scale_y_discrete(breaks = brks,   # Breaks
         labels = lbs) +
  scale_x_continuous(breaks = brksx,   # Breaks
         labels = lbsx) + 
  theme(axis.text.y = element_text(size = 6)) +
  geom_tile(colour = "white") +
  facet_grid(rows = vars(Issue_Year)) + 
  coord_cartesian(xlim = c(1, 31)) +
  scale_fill_gradient(low="white", high="red") + 
  labs(x = 'Month Day',
        y = 'Month') +
  guides(fill = guide_colorbar(
  title = 'Ammount of Tickets'))

```


####b.Number of tickets during day hours for the top 7 most common vilation types
```{r message=FALSE, warning=FALSE}
brks <- seq(1, 24, 1)
labs <- seq(1, 24, 1)
LA_sql  %>% 
  filter(Violation_Description == 'NO PARK/STREET CLEAN' |
         Violation_Description == 'METER EXP.' |
         Violation_Description == 'RED ZONE' |
         Violation_Description == 'PREFERENTIAL PARKING' |  
         Violation_Description == 'DISPLAY OF TABS' |  
         Violation_Description == 'NO PARKING' |
         Violation_Description == 'DISPLAY OF PLATES') %>%
  group_by(Violation_Description, Issue_Hour) %>%
  filter(is.na(Issue_Hour) == F) %>% 
  summarise(n = n()) %>%
  ggplot(aes(x = Issue_Hour,
             y = n)) + 
  geom_line(aes(group = Violation_Description,
                color = Violation_Description)) +
  scale_x_continuous(breaks = brks,   # Breaks
         labels = labs) +
  labs(x = 'Day Hour',
        y = 'Ammount of Tickets',
       title = 'Number of Ticket during Day Hours') +
  guides(fill = guide_colorbar(
  title = 'Violation Type'))

  

```












