---
title: "Biostat M280 Homework 3"
subtitle: Due Mar 1 @ 11:59PM
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use tidyverse and bash to explore following two data sets.

## Q1 LA City Employee Payroll

The `/home/m280-data/la_payroll/City_Employee_Payroll.csv` file on teaching server contains payroll information of LA City employees in years 2013-2018. It was downloaded from [LA City Controller's Office](https://controllerdata.lacity.org/Payroll/City-Employee-Payroll/pazn-qyym). Make a Shiny app to facilitate visualization of this data. 

1. For efficiency of the Shiny app, you should first pre-process, pare down, tidy, and save the data, e.g., as a compressed RDS file, to be used in the app.

```{bash}
ln -sf /home/m280data/la_payroll/City_Employee_Payroll.csv LAEmployeePayroll.csv
ls -l NYParkingViolations.csv
```

####A.Load packages and read in cvs file
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readr)
library(data.table)
library(dplyr)
library(rio)
library(magrittr)
library(lubridate)
city_ep <- read_csv('LAEmployeePayroll.csv') #city employee and payroll
colnames(city_ep)
```

####B.Process colnames and chose variables needed then export tibble in a rds file.

```{r}
names(city_ep) <- str_replace_all(names(city_ep), ".\\(.*\\)", "") 
#remove parenthesse
names(city_ep) <- str_replace_all(names(city_ep), " ", "_")
library("magrittr")
city_ep %<>%
  select(Department_Title, Job_Class_Title,
        Year, Total_Payments, Base_Pay,
        Overtime_Pay, Other_Pay,
        Average_Benefit_Cost)
##save as rds
library(rio)
export(city_ep, "LAep/LAep.rds")
```

0. Visualize any other information you are interested in.

I'd like to see the 6-year average payments composition of a chosen department.

0. Publish your Shiny app to <https://www.shinyapps.io> and share the link.

See my Shiny app on <https://suedez.shinyapps.io/laep/>.

## Q2 LA City Parking War

The SQLite database `/home/m280data/la_parking/LA_Parking_Citations.sqlite` on teaching server contains information about parking tickets in LA City. It was downloaded from [LA Open Data Portal](https://data.lacity.org/A-Well-Run-City/Parking-Citations/wjz9-h9np). Connect to the database and answer following questions using plots and summary statistics. In this exercise, you are **not** allowed to load whole data into memory. Use the _transform in database, plot in R_ strategy.

1. How many tickets are in this data set? Which time period do these tickets span? Which years have most data?

```{r}
library("DBI")
library("RSQLite")
```


```{bash}
ln -sf /home/m280data/la_parking/LA_Parking_Citations.sqlite LAparking.sqlite
ls -l LAparking.sqlite
```

```{r message=FALSE, warning=FALSE}
if (Sys.info()[["sysname"]] == "Linux") {
  db <- dbConnect(RSQLite::SQLite(), 
                  dbname = "LAparking.sqlite")
} else if (Sys.info()[["sysname"]] == "Darwin") {
  db <- dbConnect(RSQLite::SQLite(), 
                  dbname = "./LA_Parking_Citations.sqlite")
                  }
dbListTables(db)
LA_sql <- dplyr::tbl(db, "latix") #use tbl function in dplyr package
str(LA_sql) #????????????????????????????????????
```

```{r message=FALSE, warning=FALSE, paged.print=TRUE}
LA_sql %>% print(width = Inf)
LA_sql %>% summarise(n = n()) -> ntickets
library(lubridate)

##time range
options(scipen=100) # don't round
LA_sql %>% mutate(time = Issue_Year * 10e7 + 
                    Issue_Month * 10e5 + 
                    Issue_Day * 10e3 +
                    Issue_Hour * 10e1 + 
                    Issue_Minute) %>%
  summarise(Time_begin = min(time),
             Time_end = max(time))
##wich year most tickets
LA_sql %>% group_by(Issue_Year) %>% summarise(n = n()) %>% 
  ggplot() +
  geom_bar(stat = 'identity', aes(x = Issue_Year,
                                  y = n)) +
  labs(title = paste('Ammounts of tickets in year 2015-2019'), 
        x = 'Issue Year',
        y = 'Ammout of Tickets')
```

    As it shown above, the time range is from 1:00,July 2nd, 2015 to 23:58, January 25th, 2019. The year has most tickets is 2017.

0. When (which hour, weekday, month day, and month) are you most likely to get a ticket and when are you least likely to get a ticket?

```{r}
LA_sql %>% group_by(Issue_Month) %>% summarise(n = n()) %>% 
  arrange(desc(n)) -> MONTH
LA_sql %>% group_by(Issue_Day) %>% summarise(n = n()) %>% 
  arrange(desc(n)) -> DAY
LA_sql %>% group_by(Issue_Wday) %>% summarise(n = n()) %>% 
  arrange(desc(n)) -> WDAY
LA_sql %>% group_by(Issue_Hour) %>% summarise(n = n()) %>% 
  arrange(desc(n)) -> HOUR

```

    As most likely month, month day, week day, hour to get a ticket are, respectively, August, 13th, Wednesday，12:00.
    The least likely month, month day, week day, hour to get a ticket are, respectively, February, 31th, Monday, 5:00. But I won't be lax at the 31th since not every month has the 31th so that 
    
0. Which car makes received most citations?

```{r}
LA_sql %>% group_by(Make) %>% summarise(n = n()) %>% 
  arrange(desc(n)) %>% filter(is.na(Make) == F)

```

    TOYT received most citations.

0. How many different colors of cars were ticketed? Which color attracted most tickets?

```{r}
LA_sql %>% group_by(Color) %>% summarise(n = n()) %>% arrange(desc(n)) %>%
  filter(is.na(Color) == F) 
```

    The color black is the most attractive color, but maybe it is because black is the most common color for cars. There are 103 different colors.

0. What are the most common ticket types?

```{r}
LA_sql %>% group_by(Violation_Description) %>% summarise(n = n()) %>% arrange(desc(n))
```

    The most common ticket type is No Park/Street Clean.

0. How much money was collected on parking tickets in 2016, 2017 and 2018?

```{r}
LA_sql %>% filter(Issue_Year == 2016 | Issue_Year == 2017 |
                  Issue_Year == 2018) %>% select(Fine_amount, 
                                           Issue_Year) %>%
  group_by(Issue_Year) %>% 
  summarise(Money = sum(Fine_amount)) %>%
  ggplot() + geom_bar(stat = 'identity',
                      aes(x = Issue_Year, 
                          y = Money, fill = Money)) +
  labs(title = 'Ammounts of fine collected in year 2016-2018', 
        x = 'Year',
        y = 'Ammout of Fine')
```

0. If you've been ticketed in LA County, did you find your ticket in this data set?

    I don't even have a car hahaha.


0. Read the blog <http://www.brettrics.com/9-million-parking-tickets-la/> and try to reproduce plots using your data.

```{r}
LA_sql %>% 
```











